Sampling Strategies in Decisions from Experience

### 2022-02-26 manuscript.Rmd
Update Results

```{r packages}
# load required packages
pacman::p_load(papaja,
               readr,
               tidyr, 
               dplyr,
               ggplot2,
               viridis,
               scico,
               latex2exp)
```

@@

```{r w-shapes, fig.cap="(ref:weighting)"}

@@

# plot
ggplot(wf, aes(p, wp, group = gamma)) +
  facet_wrap(~delta, labeller = as_labeller(labeller_delta, default = label_parsed)) + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x = expression(p),
       y = expression(w(p)), 
       color = expression(gamma)) + 
  theme_apa() + 
  geom_line(aes(color = gamma)) +
  scale_color_scico(palette = "berlin") + 
  geom_abline(intercept = 0, slope = 1, color = "gray", linetype = "dashed", size = 1) # identity line
```

@@

```{r}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices_p <- read_csv("data/choices/choices.csv", col_types = cols) # piecewise integration

# Choices

# determine normative choice objective and experienced
fr_rates_p <- choices_p %>%
  mutate(norm = case_when(ev_ratio > 1 ~ "A", ev_ratio < 1 ~ "B"), # normative choice objective EV
         ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2),
         norm_exp = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) # normative choice experienced EV / mean

# objective
fr_rates_p_o <- fr_rates_p %>%
  filter(!is.na(norm)) %>%
  group_by(s, boundary, a, rare, norm, choice) %>%
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2), # false response rates
         type = case_when(norm == "A" & choice == "B" ~ "Safe", norm == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>% # remove correct responses
  select(-c(norm, n)) %>% 
  mutate(norm = "Latent")

# experienced
fr_rates_p_e <- fr_rates_p %>%
  filter(!is.na(norm_exp)) %>%
  group_by(s, boundary, a, rare, norm_exp, choice) %>%
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2), # false response rates
         type = case_when(norm_exp == "A" & choice == "B" ~ "Safe", norm_exp == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>%  # remove correct responses
  select(-c(norm_exp, n)) %>% 
  mutate(norm = "Sample")

# merge data frame
fr_rates_p <- bind_rows(fr_rates_p_o, fr_rates_p_e) %>% 
  mutate(rare = case_when(rare == "none" ~ "Middle",
                          rare == "attractive" ~ "High", 
                          rare == "unattractive" ~ "Low"))

labeller_a <- function(string) {
  TeX(paste("$\\a=$", string, sep = ""))  
}

fr_rates_p %>% 
  ggplot(aes(s, rate)) +
  geom_point(aes(color = norm, shape = type)) +
  facet_grid(rare~a, labeller = labeller(rare = label_value, a = as_labeller(labeller_a, default = label_parsed))) + 
  scale_color_scico_d(palette = "berlin")+ 
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1)) + 
  scale_y_continuous(limits = c(0, 1), breaks = c(0, .5, 1)) +
  labs(x = "s",
       y = "False Response Rate", 
       color = "Normative\nEV",
       shape = "False\nResponse") +
  theme_apa()
```

@@

<!-- Results -->

>>>>

```{r packages}
# load required packages
pacman::p_load(papaja,
               readr,
               tidyr, 
               dplyr,
               ggplot2,
               viridis,
               ggpubr,
               latex2exp)
```

@@

```{r w-shapes, fig.cap="(ref:weighting)", fig.height=7, fig.width=10}

@@

# plot
ggplot(wf, aes(p, wp, group = gamma)) +
  facet_wrap(~delta, labeller = as_labeller(labeller_delta, default = label_parsed)) + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x = expression(p),
       y = expression(w(p)), 
       color = expression(gamma)) + 
  theme_apa() + 
  geom_line(aes(color = gamma)) +
  scale_color_viridis(option = "plasma") + 
  geom_abline(intercept = 0, slope = 1, color = "gray", linetype = "dashed") # identity line
```

@@

(ref:fr-rates-p) Effect of Switching Probability on Rates of False Risky and False Safe Choices for 2-Outcome Prospects with High Rank Outcomes of Different Probability

```{r fig.cap="(ref:fr-rates-p)", fig.height=7, fig.width=10}
cols <- list(.default = col_double(),
             boundary = col_factor(),
             gamble = col_factor(),
             rare = col_factor(),
             agent = col_factor(),
             choice = col_factor())
choices_p <- read_csv("data/choices/choices.csv", col_types = cols) # piecewise integration

# false response (fr) rates

## determine normative choice according to latent and sampled EV
fr_rates_p <- choices_p %>%
  mutate(norm = case_when(ev_ratio > 1 ~ "A", ev_ratio < 1 ~ "B"), # normative choice latent EV
         ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2),
         norm_exp = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) # normative choice sampled EV

# fr rates latent EV
fr_rates_p_o <- fr_rates_p %>%
  filter(!is.na(norm)) %>% # exclude choices under EV equality
  group_by(s, boundary, a, rare, norm, choice) %>% # to distinguish type of rare event and type of false response
  summarise(n = n()) %>% # determine absolute number of responses per category
  mutate(rate = round(n/sum(n), 2), # determine correct and false response rates
         type = case_when(norm == "A" & choice == "B" ~ "Safe", norm == "B" & choice == "A" ~ "Risky")) %>% # type false response
  ungroup() %>%
  filter(!is.na(type)) %>% # remove correct response rates
  select(-c(norm, choice, n)) %>%
  mutate(norm_ev = "Latent")

# fr rates sampled EV
fr_rates_p_e <- fr_rates_p %>%
  filter(!is.na(norm_exp)) %>%
  group_by(s, boundary, a, rare, norm_exp, choice) %>%
  summarise(n = n()) %>%
  mutate(rate = round(n/sum(n), 2),
         type = case_when(norm_exp == "A" & choice == "B" ~ "Safe", norm_exp == "B" & choice == "A" ~ "Risky")) %>%
  ungroup() %>%
  filter(!is.na(type)) %>%
  select(-c(norm_exp, choice, n)) %>%
  mutate(norm_ev = "Sample")

# merge data frame
fr_rates_p <- bind_rows(fr_rates_p_o, fr_rates_p_e) %>%
  mutate(rare = case_when(rare == "none" ~ "\\in \\[.2,.8\\]",
                          rare == "attractive" ~ "\\in (0,.2)",
                          rare == "unattractive" ~ "\\in (.8,1)")) # Change fact labels

# labeller function for LaTeX math expressions in facet labels
labeller_a <- function(string) {
  TeX(paste("$\\a=$", string, sep = ""))
}

labeller_rare <- function(string) {
  TeX(paste("$\\p_{High}$", string, sep = ""))
}

ggplot(fr_rates_p, aes(s, rate)) +
  facet_grid(rare~a, labeller = labeller(rare = as_labeller(labeller_rare, default = label_parsed), a = as_labeller(labeller_a, default = label_parsed))) +
  labs(x = "s",
       y = "False Response Rate",
       color = "Normative\nEV",
       shape = "False\nResponse") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, .5)) +
  labs(x = "Switching Probability",
       y = "False Response Rate",
       color = "Normative\nExpected\nValue",
       shape = "False\nResponse\nType") +
  geom_point(aes(color = norm_ev, shape = type)) +
  scale_shape_manual(values = c(4, 16)) +
  scale_color_viridis_d(option = "plasma", end = .7) +
  theme_apa()
```

@@

(ref:weighting-parameters) Parameter Estimates of the Weighting Function of @goldsteinExpressionTheoryPreference1987 for Different Simulation Parameters

```{r fig.cap="(ref:weighting-parameters)", fig.height=7, fig.width=10}

# load data

cols <- list(.default = col_double(),
             boundary = col_factor(),
             a = col_factor(),
             parameter = col_factor())
cpt_p <- read_csv("data/estimates/estimates_cpt.csv", col_types = cols) # piecewise integration
cpt_p <- cpt_p %>% mutate(boundary = if_else(boundary == "absolute", "Absolute", "Relative"))

gamma <- cpt_p %>%
  filter(parameter == "gamma") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  labs(x = "Switching Probability",
       y = expression(gamma),
       color = "Switching\nProbability",
       shape = "Boundary") +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(labeller_a, default = label_parsed))) + 
  theme_apa() + 
  theme(axis.title.x = element_blank())

delta <- cpt_p %>%
  filter(parameter == "delta") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  facet_wrap(~a, nrow = 1) + 
  labs(x = "Switching Probability",
       y = expression(delta),
       color = "Switching\nProbability",
       shape = "Boundary") +
  theme_apa() + 
  theme(strip.text.x = element_blank())

ggarrange(gamma, delta, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

(ref:value-choice-parameters) Parameter Estimates of the Value Function of @tverskyAdvancesProspectTheory1992 and of the Logit Choice Rule for Different Simulation Parameters

```{r fig.cap="(ref:value-choice-parameters)", fig.height=7, fig.width=10}

alpha <- cpt_p %>%
  filter(parameter == "alpha") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0,1,.5)) +
  labs(x = "Switching Probability",
       y = expression(alpha),
       color = "Switching\nProbability",
       shape = "Boundary") +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(labeller_a, default = label_parsed))) + 
  theme_apa() + 
  theme(axis.title.x = element_blank())

rho <- cpt_p %>%
  filter(parameter == "rho") %>% 
  ggplot(aes(s, mean, color = s, shape = boundary)) +
  geom_point() +
  geom_pointrange(aes(ymin=`2.5%`, ymax=`97.5%`)) + 
  scale_color_viridis(option = "plasma") +
  scale_shape_manual(values = c(4, 16)) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1,.5)) +
  labs(x = "Switching Probability",
       y = expression(rho),
       color = "Switching\nProbability",
       shape = "Boundary") +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(labeller_a, default = label_parsed))) + 
  theme_apa() + 
  theme(strip.text.x = element_blank())

ggarrange(alpha, rho, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```

(ref:shapes) Estimated Graphical Shapes of the Weighting and Value Function for Different Simulation Parameters

```{r fig.cap="(ref:shapes)", fig.height=7, fig.width=10}

## weighting function

wf_p <- cpt_p %>%
  select(s, boundary, a, parameter, mean) %>%
  pivot_wider(names_from = parameter, values_from = mean) %>%
  select(-c(alpha, rho)) %>%
  expand_grid(p = seq(0, 1, .05)) %>%
  mutate(w = round(  (delta * p^gamma)/ ((delta * p^gamma)+(1-p)^gamma), 2))

wf_p_a <- wf_p %>%
  filter(boundary == "Absolute")
wf_p_r <- wf_p %>%
  filter(boundary == "Relative")
weight <- ggplot(wf_p_a, aes(x = p, y = w, group = s, color = s)) +
  geom_line(size = .5) +
  geom_line(data = wf_p_r, size = .5) + 
  scale_color_viridis(option = "plasma") +
  scale_x_continuous(breaks = seq(0, 1, .5)) +
  scale_y_continuous(breaks = seq(0, 1, .5)) +
  labs(x = "Experienced Probability",
       y = "Decision Weight",
       color = "Switching\nProbability") +
  facet_wrap(~a, nrow = 1, labeller = labeller(a = as_labeller(labeller_a, default = label_parsed))) + 
  theme_apa() + 
  theme(strip.text.x = element_blank())

vf_p <- cpt_p %>%
  select(s, boundary, a, parameter, mean) %>%
  pivot_wider(names_from = parameter, values_from = mean) %>%
  select(-c(gamma, delta, rho)) %>%
  expand_grid(x = seq(0, 20, .5)) %>%  
  mutate(v = round(x^alpha, 2))

vf_p_a <- vf_p %>%
  filter(boundary == "Absolute")
vf_p_r <- vf_p %>%
  filter(boundary == "Relative")
value <- ggplot(vf_p_a, aes(x = x, y = v, group = s, color = s)) +
  geom_line(size = .5) +
  geom_line(data = vf_p_r, size = .5) + 
  scale_color_viridis(option = "plasma") +
  scale_x_continuous(breaks = seq(0, 20, 10)) +
  scale_y_continuous(breaks = seq(0, 20, 10)) +
  labs(x = "Objective Outcome",
       y = "Subjective Value",
       color="Switching\nProbability") +
  facet_wrap(~a, nrow = 1)+ 
  theme_apa() + 
  theme(strip.text.x = element_blank())

ggarrange(weight, value, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right", labels = "AUTO") 
```
